---
title: "Assignment 2 Week 4"
author: "VibhorVyas"
date: "November 6, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)

```

## Assignment 2 Week 4 

```{r}
plot <- function(names, totals, columns, main, colors){
  colnames(totals) <- names
  
  par(las=2,mar=c(6,4,1,1))
  barplot(totals, col=colors,main=main,cex.names  = 0.6,cex.axis = 0.6)
  legend("topright", columns,fill=colors,bty = "n")
}

```
## Reading Data

```{r}
repdataFdataFStormData_csv <- read_csv("C:/Workspace/CWorkspace/learning/Coursera/CourseraAssignements/ReproducableResearch/repdataFdataFStormData.csv.bz2")
#View(repdataFdataFStormData_csv)
colnames(repdataFdataFStormData_csv)
event_types <- as.data.frame(table(repdataFdataFStormData_csv$EVTYPE))
event_types <- event_types[order(event_types$Var1), ]
## Separate data for Health and property so that those can be read and randered as plots after data formatting.

health <- repdataFdataFStormData_csv[,(c(8,23:24))]
property<-repdataFdataFStormData_csv[,c(8,25:28)]

```

## Including Plots
Formatting the data and removing NA values
```{r}
table(property$PROPDMGEXP)
table(property$CROPDMGEXP)

property$PROPDMGEXP<-factor(property$PROPDMGEXP,levels=c("H","K","M","B","h","m","O"))
property$PROPDMGEXP[is.na(property$PROPDMGEXP)] <- "O"

property$CROPDMGEXP<-factor(property$CROPDMGEXP,levels=c("K","M","B","k","m","O"))
property$CROPDMGEXP[is.na(property$CROPDMGEXP)] <- "O"

property$PROPDMGEXP <- as.character(property$PROPDMGEXP)
property$CROPDMGEXP <- as.character(property$CROPDMGEXP)

```
# Converting the values of Thousands, Millions and Billions. 

```{r}

property$PROPDMGMLT <- 0
property$CROPDMGMLT <- 0

property$PROPDMGMLT[grepl("h", property$PROPDMGEXP,ignore.case = TRUE)]<-100
property$PROPDMGMLT[grepl("k", property$PROPDMGEXP,ignore.case = TRUE)]<-1000
property$PROPDMGMLT[grepl("m", property$PROPDMGEXP,ignore.case = TRUE)]<-1000000
property$PROPDMGMLT[grepl("b", property$PROPDMGEXP,ignore.case = TRUE)]<-1000000000
property$PROPDMGMLT[grepl("o", property$PROPDMGEXP,ignore.case = TRUE)]<-1

property$CROPDMGMLT[grepl("k", property$CROPDMGEXP,ignore.case = TRUE)]<-1000
property$CROPDMGMLT[grepl("m", property$CROPDMGEXP,ignore.case = TRUE)]<-1000000
property$CROPDMGMLT[grepl("b", property$CROPDMGEXP,ignore.case = TRUE)]<-1000000000
property$CROPDMGMLT[grepl("o", property$CROPDMGEXP,ignore.case = TRUE)]<-1


```
Finding the property datamage and crop damage data. Also finding total daamage

```{r}
property$PROPDMG <- property$PROPDMG * property$PROPDMGMLT
property$CROPDMG <- property$CROPDMG * property$CROPDMGMLT
property$total <- property$PROPDMG + property$CROPDMG

```
Ditermin health impact by finding Fatalities and Injuries

```{r}

health.totals <- aggregate(cbind(FATALITIES,INJURIES) ~ EVTYPE, data = health, sum, na.rm=TRUE)
health.totals$TOTAL <- health.totals$FATALITIES + health.totals$INJURIES
health.totals <- health.totals[order(-health.totals$TOTAL), ]
health.totals <- health.totals[1:25,]


```
Plotting the graph of health totals and Fatalities and Injuris
```{r}
plot(health.totals$EVTYPE,
     as.matrix(t(health.totals[,c(-1,-4)])),
     colors = c("blue","red"),
     columns = c("Fatalities","Injuries"),
     main = "Most Harmful Event Type in the United States")

```
## from the above graph its clear indicator that Tornedo is causing Maximum Injuries indicated in Red) and Fetalities (indicated in Blue)

Finding economic impact 

```{r}
economic.total <- aggregate(cbind(PROPDMG,CROPDMG, total) ~ EVTYPE, data = property, sum, na.rm=TRUE)
economic.crop <- economic.total[order(-economic.total$CROPDMG), ]
economic.crop <- economic.crop[1:25,]

economic.prop <- economic.total[order(-economic.total$PROPDMG), ]
economic.prop <- economic.prop[1:25,]


```
plotting the graphs of corp damage and property damage due to weather

```{r}
plot(economic.prop$EVTYPE,
     as.matrix(t(economic.prop[,c(-1,-3,-4)])),
     colors = c("blue","red"),
     columns = c("Property Damage"),
     main = "Economic Impact - Property Damage")

plot(economic.crop$EVTYPE,
     as.matrix(t(economic.crop[,c(-1,-2,-4)])),
     colors = c("blue","red"),
     columns = c("Crop Damage"),
     main = "Economic Impact - Crop Damage")

```
## From above graphs its clearly identified Draught is major cause of Crop damage and Flood is major cause of Property Damage.